\section{Related works} \label{s:related}
\textcolor{blue}{ In this section, we briefly discuss several recent studies on the mortality prediction of ICU patients and analyze their main differences from that approach that we adopted in this study. We provided a more comprehensive review for the interested readers in the supplementary materials.}

\textcolor{blue}{Calvert \textit{et al.} developed an algorithm, namely, \textit{AutoTriage}, which uses eight common clinical variables, mostly from physiological measurements, and 2/3 discretized variables \cite{Calvert2016}. These clinical variables produce sub-scores. A combination of weighted sub-scores is used as the final score for the mortality prediction of ICU patients. They conducted experiments on MIMIC-III (Medical Information Mart for Intensive Care III) hospital database of ICU patients. Bhattacharya \textit{et al.} conducted a study for ICU mortality prediction addressing the class imbalance issue of the ICU data in the context of binary classification through a feature transformation approach \cite{Bhattacharya2017}. They used demographic data, 37 lab investigations and some physiological signal measurements for their experiments. Xie \textit{et al.} outlined the essential procedures and concepts for developing prediction models for in-hospital mortality prediction for ICU patients in \cite{Xie2017}. In their work, Artificial Neural Network (ANN), Decision Tree (DT), and Support Vector Machines (SVM) exhibited promising results in analyzing large and heterogeneous data compared to Logistic Regression (LR). They used physiological variables along with others. Awad \textit{et al.} proposed a method called Early Mortality Prediction for ICU patients (EMPICU) \cite{Awad2017} to predict mortality after six hours from admission to ICU. They included demographic, physiological, vital signs as well as laboratory test variables collected from the MIMIC-II database. Nguyen \textit{et al.} proposed a deep learning architecture based on Long Short-Term Memory networks (LSTM) with layered attentions mechanism to predict ICU mortality to address the issue of missing measurements \cite{Nguyen2017}. They used 41 measure types including vital signs.} 

\textcolor{blue}{In an attempt to present benchmarks, Johnson \textit{et al.} conducted 38 experiments of 28 published studies that used the MIMIC database to reproduce the cohorts (i.e., a group of variables) used in these studies in the context of the performance of mortality prediction models \cite{Johnson2nd2017}. Darabi \textit{et al.} developed a technique in which they applied Gradient Boosting Decision Trees (GBT) and deep neural networks to predict the mortality of ICU patients \cite{Darabi2018}. They used grid-search to find the optimal parameters for the model. They conducted their experiments on the medical codes (diagnosis codes, procedure codes, diagnosis-related group code) collected from the MIMIC-III database.} 

\textcolor{blue}{Sadeghia \textit{et al.} proposed a method which utilized statistical and signal-based features for early hospital mortality prediction \cite{Sadeghi2018}. They used only vital signals, i.e., heart signals of the patients and these were collected from the MIMIC-III database. Zheng and Shi utilized LSTM and Recurrent Neural Network (RNN) based deep learning techniques for ICU mortality prediction \cite{Zheng2018}. They used a statistical approach to preprocess the data. They presented a data imputation method based on the Gaussian process. They used ICU data containing 36 variables from the PhysioNet. These 36 features are mentioned in terms of importance from a medical perspective.} 

\textcolor{blue}{Zahid and Joon Lee explored deep learning techniques focusing on the Self Normalizing Neural network (SNN) for predicting mortality of the ICU patients \cite{Zahid2018}. They used demographic information, vital physiological signs, progress notes by physicians and nurses, reports from imaging studies, lab test results, International Classification of Diseases-9 (ICD-9) codes, daily Simplified Acute Physiology Score (SAPS) and Sequential Organ Failure Assessment (SOFA) score, discharge summaries, hospital lengths of stay (LoS), and output of the hospital mortality. Purushotham \textit{et al.} performed a study for presenting benchmark results for clinical prediction tasks (e.g., mortality prediction, LoS prediction, and ICD-9 code group prediction) using deep learning models, Super ICU Learner Algorithm (SICULA), and SAPS-II and SOFA \cite{Purushotham2018}. They found the deep learning models performing consistently better than all other approaches.} 

\textcolor{blue}{Gennatas \textit{et al.} proposed Expert-Augmented Machine Learning (EAML) that guides the extraction of expert knowledge and it's integration into machine learning models \cite{Gennatas2019}. They conducted experiments on MIMIC-II \& MIMIC-III datasets for predicting mortality of the ICU patients. They used vital signs along with other clinical variables in their experiments. Caicedo-Torres and Gutierrez proposed a deep multi-scale convolutional architecture trained on the MIMIC-III dataset for predicting mortality of ICU patients \cite{Torres2019}. They demonstrated visual explanations focusing on how the network treats these inputs as important features. They used 22 different variables including vital signs which are somewhat similar to the concepts used by the SAPS-II score. De Lange \textit{et al.} proposed a cumulative prognostic score model for predicting mortality of ICU patients who are older than 80 years \cite{Lange2019}. They developed a multi-variable LR model in which variables are selected using LASSO. Data of 306 ICU patient from 24 European countries were collected and their 24 clinical variables were used in the experiment.} Herland \textit{et al.} conducted a comprehensive survey on clinical data mining applications based on big data in health informatics \cite{Herland}. Cai \textit{et al.}  proposed an approach based on the Bayesian network to develop models using EHRs for real-time prediction of several targets, including the length of hospital stay, mortality, and readmission of hospitalized patients \cite{Cai:2016}.

Now we analyze how our study is fundamentally different from the research endeavors mentioned above. To predict the survival of ICU patients, most of the above studies used vital signs or demographic variables and some focused mainly on clinical notes, whereas we only used lab test results in our approach. Some works indeed consider lab tests measurements albeit only a small set thereof and that too in combination with other features. On the contrary, we consider only lab tests measurements and a large set thereof. Next, we find the majority of the above studies focusing on a single machine learning/data mining technique. However, our work proposes a hybrid approach: features are selected at first and then ensemble models are generated using the FVC technique with various ensemble classification techniques. Therefore we believe that our current approach applies to any CDS related systems in general. Besides, a great advantage of our FVC based approach, compared to any machine learning technique, is its inherent capability of tackling the curse of dimensionality\cite{Meyer2018} of hospital database Big data. Furthermore, some of the previous studies deal with particular disease-oriented data. But we generally work for the mortality of ICU patients irrespective of any disease. Also, we have used different kinds of datasets for the experiments. More will be elaborated at later sections of this paper.


%\textcolor{blue}{Additionally, some of the above deep learning methodology based studies deal with Big data in the context of in-hospital mortality of ICU patients. According to Meyer \textit{et al.} \cite{Meyer2018} adding more variables into the actual data matrix increases the dimension of the matrix which may become computationally burdensome for this kind of model. But FVC technique \cite{mehedy-masud:2017:fvc} is reverse of this, deals with large data and has shown the improved performance on training or testing time.}

Finally, we would like to mention that, a subset of the authors were involved in proposing the FVC approach in \cite{mehedy-masud:2017:fvc} and ORCU (a novel orthogonal clustering technique) in \cite{mehedy-masud:2018:frmwrk} where some preliminary experiments and findings were discussed. In this study, we proposed further improvements of \cite{mehedy-masud:2017:fvc,mehedy-masud:2018:frmwrk} and conducted extensive experiments with real clinical data thereby reaching a promising milestone in this research endeavor.